---
title: "in-class_ex10"
author: "Su Yiin Ang"
date: "3/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE,
                      fig.retina = 3,
                      warning = FALSE,
                      eval = TRUE)
```


## Installing and launching R packages

```{r}
packages = c('tmap','sf','tidyverse')

for(p in packages){
  if(!require(p, character.only = T)){
  install.packages(p)
    }
  library(p, character.only = T)
  }
```

## Import data

```{r}
sgpools <- read_csv('data/aspatial/SGPools_svy21.csv')
```


## combine xcoord and ycoord
if you have lat and long, lat is xcoord & long is ycoord
```{r}
sgpools_sf <- st_as_sf(sgpools, 
                       coords = c('XCOORD', 'YCOORD'),
                       crs = 3414) #sg epsg code
```

# point a point symbol map

```{r}
tmap_mode('view')
tm_shape(sgpools_sf) + #base layer
tm_bubbles(col='red',
          size = .5,
          border.col = 'black',
          border.lwd = 1)

```
## Proportional symbol map

```{r}
tm_shape(sgpools_sf) +
  tm_bubbles(col = 'red',
            size = 'Gp1Gp2 Winnings',
            border.col = 'black',
            border.lwd = 0.5) #control width of line.

```
## Proportional symbol map - colour by outlet type

```{r}
tm_shape(sgpools_sf) + 
tm_bubbles(col='OUTLET TYPE',
          size = 'Gp1Gp2 Winnings',
          border.col = 'black',
          border.lwd = 0.5)
```

## create two views and synchronise them

```{r, echo = FALSE}
tm_shape(sgpools_sf) + 
  tm_bubbles(col='OUTLET TYPE',
          size = 'Gp1Gp2 Winnings',
          border.col = 'black',
          border.lwd = 0.5) +
  tm_facets(by = 'OUTLET TYPE',
            nrow = 1,
            sync = TRUE)
```

# Choropleth mapping

## import data

```{r}
mpsz <- st_read(dsn = 'data/geospatial',
                layer = 'MP14_SUBZONE_WEB_PL') #shape file 
```

## bring in population data

```{r}
popagsex <- read_csv('data/aspatial/respopagsex2000to2018.csv')

```

## data wrangling


### with pivot wider - should use this one

```{r}
popagsex %>%
  filter(Sex == 'Males') %>%
  filter(Time == 2018) %>%
  pivot_wider(names_from = AG, values_from = Pop)
```

### with spread

```{r}
popagsex2018_male <- popagsex %>%
  filter(Sex == 'Males') %>%
  filter(Time == 2018) %>%
  spread(AG, Pop) %>% #try with pivot wide
  mutate(YOUNG = `0_to_4` + `5_to_9` + `10_to_14` + `15_to_19` + `20_to_24` ) %>%
  mutate(`ECONOMY ACTIVE` = rowSums(.[9:13]) + rowSums(.[15:17])) %>%
  mutate(`AGED` = rowSums(.[18:22])) %>%
  mutate(`TOTAL` = rowSums(.[5:22])) %>%
  mutate(`DEPENDENCY` = (`YOUNG` + `AGED`) / `ECONOMY ACTIVE`) %>%
  mutate_at(.vars = vars(PA,SZ), .funs = funs(toupper)) %>%
  select(`PA`, `SZ`,`YOUNG`, `ECONOMY ACTIVE`, `AGED`, `TOTAL`,`DEPENDENCY`) %>% #only select these variables
  filter(`ECONOMY ACTIVE` > 0)
```

## combine with map layer

```{r}
mpsz_agemale2018 <- left_join(mpsz, popagsex2018_male,
                              by = c('SUBZONE_N' = 'SZ')) #join by common variable
```

## plot choropleth

```{r}
tmap_mode('plot')
qtm(mpsz_agemale2018,
    fill = 'DEPENDENCY')
```


```{r}
tm_shape(mpsz_agemale2018) +
  tm_polygons('DEPENDENCY')
```

```{r}
tm_shape(mpsz_agemale2018) +
  tm_fill('DEPENDENCY') +
  tm_borders(lwd = 0.1, alpha = 1)
```

## classification methods of tmap

```{r}
tm_shape(mpsz_agemale2018) +
  tm_fill('DEPENDENCY',
          n = 8, 
          style = 'quantile') + #default is prettty
  tm_borders(alpha = 0.5)
```

## colour scheme

```{r}
tm_shape(mpsz_agemale2018) +
  tm_fill('DEPENDENCY',
          n = 6,
          style = 'quantile',
          palette = 'Blues') +
  tm_borders(alpha = 0.5)
```

### inverse the colour

```{r}
tm_shape(mpsz_agemale2018) +
  tm_fill('DEPENDENCY',
          n = 6,
          style = 'quantile',
          palette = '-Blues') + #add - before colour
  tm_borders(alpha = 0.5)
```


