---
title: "In-class_Ex07"
author: "Su Yiin Ang"
date: "2/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installing and launching R packages

First, we will check if tidyverse packages have been installed. If not, R will go ahead and install tidyverse packages before launching them.

```{r, message = FALSE, echo = FALSE}
packages = c('tidyverse')

for(p in packages){
  if(!require(p, character.only = T)){
  install.packages(p)
    }
  library(p, character.only = T)
  }
```

## Importing data file

### Importing using readr package

Importing Exam_data.csv by using *read_csv()* of **readr** package.

```{r, message = FALSE}
exam <- read_csv("data/Exam_data.csv")
```

## Plotting a bar chart

```{r}
ggplot(data = exam, aes(x = RACE)) + 
  geom_bar()
```

## Plotting a dotplot

```{r}
ggplot(data = exam, aes(x = MATHS)) + 
  geom_dotplot(dotsize = 0.5, binwidth = 2.5) +
  scale_y_continuous(NULL, breaks = NULL)
```


## Plotting a histogram

```{r}
ggplot(data = exam, aes(x = MATHS)) + 
  geom_histogram(bins = 20, 
                 colour = "black",
                 fill = "light blue")
```

## Plotting a stacked bar chart

```{r}
ggplot(data = exam, aes(x = MATHS, fill = GENDER)) + 
  geom_histogram(bins = 20, 
                 colour = "grey30")
```


## Plotting a density plot

```{r}
ggplot(data = exam, aes(x = MATHS, fill = GENDER)) + 
  geom_density()
```

## Plotting a boxplot

```{r}
ggplot(data = exam, aes(x = MATHS, y = GENDER)) + 
  geom_boxplot(notch =TRUE, fill = 'light blue')
```


## Plotting boxplot and scatterplot

```{r}
ggplot(data = exam, aes(y = MATHS, x = GENDER)) + 
  geom_boxplot(fill = 'light blue') + 
  geom_point(position = "jitter",
             size = 0.5)
```


## Plotting violin plot

See this distribution to compliment the boxplot.

alpha in boxplot refers to opacity as per Tableau.

Violin plot is asymmetric.
```{r}
ggplot(data = exam, aes(y = MATHS, x = GENDER)) + 
  geom_violin(fill = 'light blue') +
  geom_boxplot(alpha = 0.5) 
```

## Plotting a scatter plot
```{r}
ggplot(data = exam, aes(x = MATHS, y = ENGLISH)) + 
  geom_point()
```


## Plotting mean value to boxplot

Plot boxplot first before summary statistics.

```{r}
ggplot(data = exam, aes(x = GENDER, y = MATHS)) + 
  geom_boxplot() +
  stat_summary(geom = "point",
               fun.y="mean",
               colour = "red",
               size =4)
```

## Plotting bst fit curve on scatterplot

default method is loess

lm = least square method

```{r}
ggplot(data = exam, aes(x = MATHS, y = ENGLISH)) + 
  geom_point() + 
  geom_smooth(size = 0.5, method = lm)
```


## Change reset axis to start at 0
```{r}
ggplot(data=exam, aes(x= MATHS, y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm,
              size=0.5) +
  coord_cartesian(xlim=c(50,100),
                  ylim=c(50,100))
```

